Persistence of marine populations under climate and fishing

### 2014-03-03 Writing/MS.tex
proofread through discussion

Understanding interactions among disturbances will help to design management for populations subjected to these disturbances. The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.  Here we have built a general model to examine how climate and harvesting interact to affect species persistence by incorporating dispersal and reproduction.

For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the climate velocity, the better a population can adjust to harvest and climate change.  More interestingly, we found a negative relationship between the critical harvesting rate and the climate velocity.  That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.  This is an indication of an interaction between the stressors.

To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.  We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.  We found similar results from the analytically derived biomass and the simulation derived biomass. This indicates that this result is robust to changes in the dispersal kernel.  We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \cite{Crainetal2008, DarlingCote2008,Nyeetal2013}.  We could also have measured the effect by the percentage drop caused by the stressor(s) and used a multiplicative null prediction for the effect of both stressors.  In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.  Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.

@@

Our results suggest that particular combinations of harvesting and climate velocity will affect some species more than others. As shown in Figure \ref{baseline}, species with a higher reproductive rate and a longer average dispersal distance will better track a high climate velocity relative to a species that has a low reproductive rate and short dispersal distance. The finding that a higher reproductive rate can sustain higher climate velocities and harvesting rates is intuitive, especially because harvesting rate and reproductive rate cancel each other out. However it is worth pointing out that a higher reproductive rate can be generated either by shorter generation times or higher fecundity. Finding that species with shorter generation times can better keep up with shifts in climate is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species \citep{Perryetal2005}. While higher reproductive rates improved a population's ability to persist, there was a tradeoff in increasing dispersal distances. At low speeds, we found that a short dispersal dispersal distance improved the maximum harvesting rate a population could sustain while at higher speeds a longer dispersal distance improved the maximum climate velocity in which the population could persist (Figure \ref{baseline}). This tradeoff is due to the proportion of dispersing  offspring at time step $t$ which lands within the patch at time step $t + 1$. When climate is shifting slowly, a large dispersal distance sends most offspring ahead of the patch, while with faster climate velocities a long dispersal distance allows the population to make it to the new patch (Figure \ref{baseline}). Thus climate velocity will selectively favor species with dispersal distances best matched to the rate of shift.

>>>>

Understanding interactions among disturbances will help to design management for populations subjected to these disturbances. The co-occurrence of climate change-driven range shifts and fishing mean that there is the potential for synergistic interactions, which have been largely unexamined.  Here we have studied
 a general model that incorporates dispersal and reproduction to examine how climate and harvesting interact in their effects on species persistence and biomass.

For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the climate velocity, the better a population can adjust to harvest and climate change.  Further, we found a negative relationship between the critical harvesting rate and the climate velocity.  That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.  This is evidence that the stressors interact since each stressor's ability to drive the population extinct depends on the severity of the other stressor.

To quantify the interaction between the stressors, we measured the synergy between their effects on population biomass.  We found positive synergy between the stressors and that the synergy is greatest in the region of parameter space where the equilibrium biomass is smallest.  We chose to measure the effect of each stressor by the absolute drop in biomass caused by the stressor, and we used the sum of the individual effects for our null prediction of the effect of both stressors, as in \cite{Crainetal2008, DarlingCote2008,Nyeetal2013}.  In general, measuring synergy against an additive null prediction is more conservative than measuring synergy multiplicatively: the presence of additive synergy implies multiplicative synergy, but not vice versa \citep{Crainetal2008, Foltetal1999}.  Since we found small levels of positive additive synergy between the two stressors, other measures of synergy might show even higher levels of interaction.

@@

Our results suggest that particular combinations of harvesting and climate velocity will affect some species more than others. Species with a higher reproductive rate and a longer average dispersal distance will better track a high climate velocity relative to a species that has a low reproductive rate and short dispersal distance (Figure \ref{baseline}). The finding that a higher reproductive rate can sustain higher climate velocities and harvesting rates is intuitive, especially because harvesting rate and reproductive rate cancel each other out. However it is worth pointing out that a higher reproductive rate can be generated either by shorter generation times or higher fecundity. Finding that species with shorter generation times can better keep up with shifts in climate is in agreement with empirical work which has found that fish which shifted in response to warming in North Sea had faster life histories than non shifting species \citep{Perryetal2005}. While higher reproductive rates improved a population's ability to persist, increasing dispersal distances did not necessarily. At low speeds, we found that a short dispersal dispersal distance improved the maximum harvesting rate a population could sustain while at higher speeds a longer dispersal distance improved the maximum climate velocity in which the population could persist (Figure \ref{baseline}). This is because when climate is shifting slowly, a large dispersal distance sends most offspring ahead of the patch, while with faster climate velocities a long dispersal distance allows the population to make it to the new patch (Figure \ref{baseline}). Thus climate velocity will selectively favor species with dispersal distances best matched to the rate of shift.
